---
title: "DSEM_LDA"
output: html_document
date: "2026-01-18"
---
## Data Preparation

```{r}
library(R2jags)
library(here)

## Data Preparation

table <- read.csv("D:/Tübingen Uni/Longitudinal Data Analysis/Replication Package/yll/yllanon.csv")

data <- table

id_var  <- "id"
time_var <- "wave"
item_vars <- c("abort","adopt","imm") 
cov_vars    <- c("gender","ethnic","edu","inc","age","state","relig") #possible covariates if we need)

# removing "dont know" and "I haven’t thought much about it" responses

data$abort[data$abort %in% c(5, 6)] <- NA

data$adopt[data$adopt %in% c(8, 9)] <- NA

data$imm[data$imm %in% c(8, 9)] <- NA

#reversing abortion and adoption variables since they point opposite the construct of conservative cultural shift

data$abort <- 5 - data$abort
data$adopt <- 8 - data$adopt


# centering data 

# Compute wave-1 means
wave1_means <- sapply(item_vars, function(v) {
  mean(data[data$wave == 1, v], na.rm = TRUE)
})

# Center all waves using wave-1 means
for (v in item_vars) {
  data[[v]] <- data[[v]] - wave1_means[v]
}

# scaling items to std.dev

item_sds <- sapply(item_vars, function(v) sd(data[[v]], na.rm = TRUE))

for (v in item_vars) {
  data[[v]] <- data[[v]] / item_sds[v]
}

# reshaping the format
 

# Unique ids and waves
ids   <- sort(unique(data[[id_var]]))
times <- sort(unique(data[[time_var]]))

N <- length(ids)
T <- length(times)
J <- length(item_vars)

# array [person, time, item]
y_data <- array(NA_real_, dim = c(N, T, J),
                dimnames = list(ids, times, item_vars))

for (j in seq_along(item_vars)) {
  tmp <- data[, c(id_var, time_var, item_vars[j])]
  names(tmp) <- c("id", "time", "y")

  wide <- reshape(tmp, idvar = "id", timevar = "time", direction = "wide")

  wide <- wide[match(ids, wide$id), , drop = FALSE]
  col_order <- paste0("y.", times)

  y_data[, , j] <- as.matrix(wide[, col_order])
}


# Save the number participants (N=552)

num_obs  <- dim(y_data)[1]

# Save the number of time points (num_time=26)

num_time <- dim(y_data)[2]

# Number of items (3)

num_items <- dim(y_data)[3]

```

```{r}
# Prepare data for one factor
data_1factor <- list(
  N = num_obs, # sample size
  Nt = num_time, # time points
  y = y_data[, , 1:3], # responses to items 1 to 3
  psi0 = diag(2), # hyperprior wishart distribution
  mu.zeta2 = rep(0, 2) # hyperprior multivariate normal distribution
)
```

```{r}
# Define the factor structure for standardization
factor_structure <- data.frame(
  indicator = 1:3,
  factor = rep(1, 3),
  loading_id = c(NA, 1, 2)  # first is fixed
)
```


################################################################################
# MODEL 1:  One-factor AR(1) with person-specific random intercept
#           and slope
################################################################################

```{r}
# Define parameters which are being estimated
params_model1 <- c(
  "alpha", "beta", "lambda.y", "nu.y", "sigma.y", "sigma.eta",
  "sigma.zeta2", "rho.zeta2", "sigma.zeta3", "sigma.delta"
)
```
# alpha: first AR(1) parameter (random intercept/baseline value)
# beta: second AR(1) parameter (slope/autoregressive coefficient)
# lambda.y: factor loadings
# lambda.y.strd: standardized factor loadings
# nu.y: factor intercepts
# sigma.y: variance of the residuals
# sigma.eta: variance of the latent factor
# sigma.zeta2: variance of the random effects
# rho.zeta2: correlation matrix  of the random effects

```{r}
# Run jags model for items 10-12 (all BAI indicators)
model1 <- jags.parallel(
  data = data_1factor, # data to use
  parameters.to.save = params_model1, # output parameters
  n.iter = 10000, # number of iterations
  n.chains = 4, # number of chains
  n.thin = 1, # number of iterations to skip
  # for avoiding autocorrelation (in this case None)
  n.burnin = 5000, # burn-in parameter
  model.file = "D:/Tübingen Uni/Longitudinal Data Analysis/Project/dsem_1.txt"
  )
```

```{r}
# Store parameter estimates
est1 <- model1$BUGSoutput$summary
# Print parameter estimates
round(est1, 2)
```

```{r}
# Print correlation matrix of the random effects
round(matrix(est1[c(
  paste0("rho.zeta2[", 1:2, ",1]"),
  paste0("rho.zeta2[", 1:2, ",2]")
), "mean"], 2, 2), 3)
```
# SOME INTERPRETATION
# -> Positive correlation between random slope and intercept which implies
#    that people with a higher starting / baseline value experience
#    a sharper increase

```{r}
# Load utility functions from GitHub code
source("D:/Tübingen Uni/Longitudinal Data Analysis/Project/standardization.R")
```

```{r}
# Calculate standardized factor loadings (see code in utils/standardization.R)
result_model1 <- run_factor_analysis_map(
  est = est1,
  factor_structure = factor_structure,
  n_factors = 1,
  person_intercept_var_name = c("sigma.zeta2[1,1]"),
  person_slope_var_name = c("sigma.zeta2[2,2]")
)
print_analysis_results(result_model1)
```

The factor loading of the first item is unrealistically high, while the third item has a very low factor loading. 

################################################################################
# MODEL 2: One-factor AR(1) with person-specific intercept + slope
#          and time-specific random slope
################################################################################

```{r}
# Define parameters which are being estimated
params_model2 <- c(
  "alpha", "beta", "lambda.y", "nu.y", "sigma.y", "sigma.eta",
  "sigma.zeta2", "rho.zeta2", "sigma.zeta3"
)
```

# alpha: first AR(1) parameter (random intercept/baseline value)
# beta: second AR(1) parameter (slope/autoregressive coefficient)
# lambda.y: factor loadings
# lambda.y.strd: standardized factor loadings
# nu.y: factor intercepts
# sigma.y: variance of the residuals
# sigma.eta: variance of the latent factor
# sigma.zeta2: variance of the person-specific random effects
# rho.zeta2: correlation matrix  of the random effects
# sigma.zeta3: variance of the time-specific random effects

```{r}
# Run jags model for items 10-12 (all BAI indicators)
model2 <- jags.parallel(
  data = data_1factor, # data to use
  parameters.to.save = params_model2, # output parameters
  n.iter = 10000, # number of iterations
  n.chains = 4, # number of chains
  n.thin = 1, # number of iterations to skip
  # for avoiding autocorrelation (in this case None)
  n.burnin = 5000, # burn-in parameter
  model.file = "D:/Tübingen Uni/Longitudinal Data Analysis/Project/dsem_2.txt"
)
```

```{r}
file.exists("D:/Tübingen Uni/Longitudinal Data Analysis/Project/dsem_model1_1factor_ar1_cross.txt")
```


```{r}
# Store parameter estimates
est2 <- model2$BUGSoutput$summary
# Print parameter estimates
round(est2, 2)
```

```{r}
# Calculate standardized factor loadings (see code in utils/standardization.R)
result_model2 <- run_factor_analysis_map(
  est = est2,
  factor_structure = factor_structure,
  n_factors = 1,
  person_intercept_var_name = c("sigma.zeta2[1,1]"),
  person_slope_var_name = c("sigma.zeta2[2,2]"),
  time_slope_var_name = c("sigma.zeta3")
)
print_analysis_results(result_model2)
```

The first factor loading is still unrealistically high. 
